Cloud correction and filtering of operational microwave humidity radiances with case specific uncertainty estimation

### 2020-09-28 Article/ici_submm_cloudcorr.tex
added description of filtering with Buehler_etal

The results show that both QRNN-single and QRNN-all are successful in predicting NFCS values for I2V, and I3V, and partially successful for I1V. The resulting error distributions are symmetric with low bias and standard deviation. For I2V and I3V, the variability of errors is even smaller than the noise. This is a consequence of predictions for cases associated with clear sky conditions. QRNN predictions are weighted mean of measurements between channels. For clear-sky conditions, the averaging of cases observed by multiple channels can cancel out the noise. The cancelling of noise is stronger in QRNN-all as all 183\,GHz channels are used concurrently. The lower prediction accuracy of I1V can be explained by the fact that it is the outermost 183\,GHz channel. The other sub-mm channels cannot provide a full coverage for the hydrometeor impact in I1V. Depending upon the atmospheric scenarios, there is always a portion of data detected by I1V, but never seen by any other channel. Though such cases are few, QRNN cannot learn to predict the clear-sky values accurately for cases which lie outside the coverage area. Besides this, lower representation of high cloud impact cases in the training set also affects QRNN accuracy.

@@

\subsection{MWI}
%
In this section, we discuss results when all 183\,GHz MWI channels are used for ``self'' cloud filtering/correction (QRNN-mwi). We also describe the performance of QRNN when sub-mm channels from ICI are used to train MWI channels.
\subsubsection{MWI alone}
%
%t
\begin{table*}[t]
	\caption{Error statistics for predictions from QRNN-mwi. The labels Filtered(7.5\,K), Predicted(All) and Predicted(7.5\,K) are same as described in Table~\ref{tab:error_statistics_ici}}
	\label{tab:statistics_mwi-alone}
	\begin{tabular}{llrr|rr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& &\multicolumn{3}{c}{QRNN-single} \\
		\cline{3-4}
		\cline{6-8}
		%		\hline
		&&   Clear-sky &   All-sky & \multicolumn{2}{c}{Filtered} & \multicolumn{2}{c}{Predicted} \\
		&&&&						(B183) & (7.5\,K)& (All) & (7.5\,K)\\
		\middlehline
		MWI-14 		&Bias     & 0.01 & -1.86 & -0.84(1.95\%) & -0.50(4.04\%) & -0.14 & -0.12 \\
					&MAE      & 0.96 &  2.59 &  1.58 &  1.26 &  1.13 &  1.00 \\
					&STD      & 1.20 &  9.02 &  3.27 &  1.96 &  1.87 &  1.53 \\
					&Skewness & 0.00 & -8.21 & -5.13 & -3.00 & -0.74 & -2.14 \\
		\middlehline
		MWI-15 		&Bias     &  0.00 & -1.24 & -0.53(1.95\%) & -0.41(2.81\%) & -0.14 & -0.12 \\
					&MAE      &  0.96 &  1.99 &  1.30 &  1.19 &  0.74 &  0.66 \\
					&STD      &  1.20 &  6.55 &  2.15 &  1.74 &  1.32 &  1.12 \\
					&Skewness & -0.01 & -9.68 & -3.67 & -2.41 & -2.42 & -2.94 \\
		\middlehline	
		MWI-16 		&Bias     &  0.00 &  -0.84 & -0.30(1.95\%) & -0.29(1.90\%) & -0.13 & -0.10 \\
					&MAE      &  0.95 &   1.63 &  1.11 &  1.10 &  0.68 &  0.62 \\
					&STD      &  1.20 &   5.39 &  1.60 &  1.54 &  1.15 &  0.96 \\
					&Skewness & -0.02 & -11.62 & -2.55 & -2.00 & -3.87 & -3.35 \\	
		\middlehline			
		MWI-17 		&Bias     &  0.01 &  -1.06 & -0.40(1.95\%) & -0.34(2.36\%) & -0.13 & -0.11 \\
					&MAE      &  0.95 &   1.83 &  1.18 &  1.13 &  0.73 &  0.67 \\
					&STD      &  1.20 &   6.29 &  1.83 &  1.61 &  1.22 &  1.03 \\
					&Skewness & -0.01 & -10.72 & -3.27 & -2.20 & -3.48 & -3.17 \\
		\middlehline			
		MWI-18 		&Bias     &  0.00 &  -0.65 & -0.22(1.95\%) & -0.25(1.39\%) & -0.11 & -0.09 \\
					&MAE      &  0.96 &   1.47 &  1.06 &  1.08 &  0.80 &  0.75 \\
					&STD      &  1.20 &   4.65 &  1.45 &  1.51 &  1.23 &  1.08 \\
					&Skewness & -0.01 & -12.84 & -1.87 & -2.09 & -2.88 & -2.40 	\\
		\bottomhline				
	\end{tabular}	
	\belowtable{} % Table Footnotes
\end{table*}
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/cloud_impact_uncertainty_MWI.pdf}	
	\caption{The average confidence intervals (2$\sigma$) plotted against the magnitude of cloud impact for all MWI channels.}
	\label{fig:uncertainty_cloud_impact}	
\end{figure}
QRNN-mwi is designed to provide basis for a cloud correction approach applicable to current satellites, that lack the higher frequency channels. Table~\ref{tab:statistics_mwi-alone} shows the results for all 183\,GHz channels of MWI. MWI channels have a higher measurement uncertainty than ICI, so a higher filtering threshold is selected. All cases with cloud correction greater than 7.5\,K are considered as cloudy and filtered out. With pure-filtering, only 4\% of the cases are filtered out for MWI-14, but the bias reduces from -1.86\,K to -0.50\,K. For the predictions, the average bias is -0.14\,K. For both filtering and correction, a significant reduction in the errors is observed but the resulting values are still high and indicate only partial success of the method. For other channels also, a similar performance is observed.  For MWI-15, cloud correction reduces the average error from -1.23\,K to -0.13\,K  and the corresponding decrease in standard deviation is 6.55\,K to 1.32\,K. Filtering the predicted cases improves the bias and standard deviation further by 15\%. Interestingly, even with partial correction, the standard deviation is lower than the measurement noise. For MWI-16, the bias for predictions is -0.12\,K compared to -0.85\,K for the measurements. The corresponding standard deviation is  1.15\,K, which is again smaller than the measurement noise(1.20\,K). MWI-17 also has a similar performance. MWI-18 is least affected by cloud impact among the five channels. The average deviation of the measurement data to NFCS values is only -0.66\,K which reduces to -0.12\,K with QRNN. After filtering 1.37\% cases with high cloud impact, only a slight improvement in the accuracy is observed but the variability of errors decreases from 1.23\,K to 1.08\,K. The results show that QRNN-mwi is partially successful in removing the cloud impact. A significant reduction in the deviation to NFCS values indicates that though most cases with high cloud impact can be removed/corrected, the cases with residual cloud impact and uncorrected cases deteriorate the performance. Using a low threshold filter can improve the accuracy, but at the cost of excessive removal of clear-sky cases.

@@

>>>>

The results show that both QRNN-single and QRNN-all are successful in predicting NFCS values for I2V, and I3V, and partially successful for I1V. The resulting error distributions are symmetric with low bias and standard deviation. For I2V and I3V, the variability of errors is even smaller than the noise. This is a consequence of predictions for cases associated with clear sky conditions. QRNN predictions are weighted mean of measurements between channels. For clear-sky conditions, the averaging of cases observed by multiple channels can cancel out the noise. The cancelling of noise is stronger in QRNN-all as all 183\,GHz channels are used concurrently. The lower prediction accuracy of I1V can be explained by the fact that it is the outermost 183\,GHz channel. Compared to other higher peaking channels, it has a higher concentration of cloudy cases or cases contaminated with surface effects. All the sub-mm channels collectively cannot provide a full coverage for the hydrometeor impact in I1V. Depending upon the atmospheric scenarios, there is always a portion of data detected by I1V, but never seen by any other channel. Though such cases are few, QRNN cannot learn to predict the clear-sky values accurately for cases which lie outside the coverage area.  Besides this, lower representation of high cloud impact cases in the training set is another factor which affects QRNN accuracy.

@@

\subsection{MWI}
%
In this section, we discuss results when all 183\,GHz MWI channels are used for ``self'' cloud filtering/correction (QRNN-mwi). QRNN-mwi is designed to provide basis for a cloud filtering/correction approach applicable to current satellites, that lack the higher frequency channels.  A comparison to an existing cloud filtering scheme described by \citet{buehler:aclou:07} is also discussed. Furthermore the we extend the prediction approach to include ICI sub-mm channels for cloud correction of MWI channels. 
\subsubsection{MWI alone}
%
\paragraph{Comparison to \citet{buehler:aclou:07}}

%t
\begin{table*}[t]
	\caption{Error statistics for predictions from QRNN-mwi. The results from filtering approach described by \citet{buehler:aclou:07} are listed under ``B183''. The labels Filtered(7.5\,K), Predicted(All) and Predicted(7.5\,K) are same as described in Table~\ref{tab:error_statistics_ici}}
	\label{tab:statistics_mwi-alone}
	\begin{tabular}{llrr|rr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& &\multicolumn{3}{c}{QRNN-mwi} \\
		\cline{3-4}
		\cline{6-8}
		%		\hline
		&&   Clear-sky &   All-sky & \multicolumn{2}{c}{Filtered} & \multicolumn{2}{c}{Predicted} \\
		&&&&						(B183) & (7.5\,K)& (All) & (7.5\,K)\\
		\middlehline
		MWI-14 		&Bias     & 0.01 & -1.86 & -0.84(1.95\%) & -0.50(4.04\%) & -0.14 & -0.12 \\
					&MAE      & 0.96 &  2.59 &  1.58 &  1.26 &  1.13 &  1.00 \\
					&STD      & 1.20 &  9.02 &  3.27 &  1.96 &  1.87 &  1.53 \\
					&Skewness & 0.00 & -8.21 & -5.13 & -3.00 & -0.74 & -2.14 \\
		\middlehline
		MWI-15 		&Bias     &  0.00 & -1.24 & -0.53(1.95\%) & -0.41(2.81\%) & -0.14 & -0.12 \\
					&MAE      &  0.96 &  1.99 &  1.30 &  1.19 &  0.74 &  0.66 \\
					&STD      &  1.20 &  6.55 &  2.15 &  1.74 &  1.32 &  1.12 \\
					&Skewness & -0.01 & -9.68 & -3.67 & -2.41 & -2.42 & -2.94 \\
		\middlehline	
		MWI-16 		&Bias     &  0.00 &  -0.84 & -0.30(1.95\%) & -0.29(1.90\%) & -0.13 & -0.10 \\
					&MAE      &  0.95 &   1.63 &  1.11 &  1.10 &  0.68 &  0.62 \\
					&STD      &  1.20 &   5.39 &  1.60 &  1.54 &  1.15 &  0.96 \\
					&Skewness & -0.02 & -11.62 & -2.55 & -2.00 & -3.87 & -3.35 \\	
		\middlehline			
		MWI-17 		&Bias     &  0.01 &  -1.06 & -0.40(1.95\%) & -0.34(2.36\%) & -0.13 & -0.11 \\
					&MAE      &  0.95 &   1.83 &  1.18 &  1.13 &  0.73 &  0.67 \\
					&STD      &  1.20 &   6.29 &  1.83 &  1.61 &  1.22 &  1.03 \\
					&Skewness & -0.01 & -10.72 & -3.27 & -2.20 & -3.48 & -3.17 \\
		\middlehline			
		MWI-18 		&Bias     &  0.00 &  -0.65 & -0.22(1.95\%) & -0.25(1.39\%) & -0.11 & -0.09 \\
					&MAE      &  0.96 &   1.47 &  1.06 &  1.08 &  0.80 &  0.75 \\
					&STD      &  1.20 &   4.65 &  1.45 &  1.51 &  1.23 &  1.08 \\
					&Skewness & -0.01 & -12.84 & -1.87 & -2.09 & -2.88 & -2.40 	\\
		\bottomhline				
	\end{tabular}	
	\belowtable{} % Table Footnotes
\end{table*}

In the study by \citet{buehler:aclou:07}, they used brightness temperature based threshold to filter out cloudy cases, using Advanced Microwave Sounding Unit B (AMSU-B). They recommend to a viewing angle dependent brightness temperature threhold at $183.31\pm1.00$ \,GHz and brightness temperature difference between $183.31\pm3.00$ and $183.31\pm 1.00$ as a measure of cloud impact. Further in the text we refer this scheme as ``B183''. MWI lacks identical 183\,GHz channels as AMSU-B, so we chose the next nearest available frequencies. We use brightness temperatures from MWI-18 (TB-MWI18) is used as proxy for $183.31\pm1.00$ \,GHz and MWI-17 for $183.31\pm3.00$ (TB-MWI17). All cases with TB-MWI18 < 235.2\,K or TB-MWI17 - TB-MWI18 < 0 are classified as cloudy. The TB-MWI18 threshold is taken from Table 1 of \citet{buehler:aclou:07} for viewing angle $44.55^\circ$. This filter is not channel specific, thus same filter is applied to MWI 183\,GHz channels. Table~\ref{tab:statistics_mwi-alone} shows a comparison of results from B183 and QRNN based cloud filtering. Due to higher measurement uncertainty in MWI than ICI, a higher filtering threshold is selected. All cases with cloud correction greater than 7.5\,K are considered as cloudy and filtered out.

The B183 filter classifies only 1.95\% of the total data as cloudy. For all five channels, rejecting these cases reduces the cloud impact significantly, but not fully.  For the outermost channel MWI-14, QRNN filter rejects almost twice as more cloudy cases than B183 and the resulting bias is almost 40\% lower. Similarly the standard deviation is almost 1.6 times lower. For MWI-15, QRNN filter removes almost 2.81\% as cloudy, and the average bias is almost 30\% lower than B183. For MWI-16 both approaches remove almost similar number of cases, and the average error and MAE are comparable. But the resulting distribution from B183 has a slightly higher spread and skewness. For MWI-17 also, QRNN filter is successful in removing more cloudy cases than B183 and results in distributions with slighly better accuracy.  However for MWI-18, the performance of B183 is better. For this channel, QRNN filter removes 28\% less cases than B183, but the overall accuracy of the two filters is not very different. The average bias after QRNN filter is -0.25\,K in comparison to -0.22\,K from B183. The corresponding difference in the standard deviation is 4\%.

Both filtering schemes are only partially successful in removing the high cloud impact. However, the results indicate a  better performance of QRNN filter than B183 for all channels except for MWI-18.
B183 filter is determined by the higher peaking channels, which when applied to other channels increases the risk of erroneous filtering. It could be possible that data in MWI-18 is cloudy but clear in MWI-14. QRNN with its channel specific approach has a clear advantage over B183. With only slighly higher rejection rate, QRNN filter can achieve data with lower bias and standard deviation. It should be noted that B183 involves AMSU-B brightness temperatures for the development of their scheme, and we do not adapt the filtering thresholds for MWI channels. This could affect the performance of B183 slightly.

\paragraph{Cloud correction and uncertainty estimation}
%
%t
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/cloud_impact_uncertainty_MWI.pdf}	
	\caption{The average confidence intervals (2$\sigma$) plotted against the magnitude of cloud impact for all MWI channels.}
	\label{fig:uncertainty_cloud_impact}	
\end{figure}

In this section, we describe the accuracy of QRNN predictions in comparison to the NFCS simulations. We also analyse the corresponding uncertainties.

Table~\ref{tab:statistics_mwi-alone} displays the different statistics quantifying the error of the best estimate. As seen the previous section, QRNN filter can only partially remove the cases with high cloud impact. With correction, a better accuracy is expected. For MWI-14, the average bias in the predicted dataset is -0.14\,K in comparison to -0.50\,K acheived by filtering alone. For other channels also, a similar performance is observed. For MWI-15, cloud correction reduces the average bias to -0.14\,K, and the corresponding standard deviation is 1.32\,K. Filtering the predicted cases improves the bias and standard deviation further by 15\%. Interestingly, even with partial correction, the standard deviation is lower than the measurement noise. For MWI-16, the bias for predictions is -0.13\,K compared to -0.84\,K for the measurements. The corresponding standard deviation is  1.15\,K, which is again smaller than the measurement noise(1.20\,K). MWI-17 also has a similar performance. MWI-18 is least affected by cloud impact among the five channels. The average deviation of the measurement data to NFCS values is only -0.65\,K which reduces to -0.11\,K with QRNN correction. After filtering 1.39\% cases with high cloud impact, only a slight improvement in the accuracy is observed but the variability of errors decreases from 1.23\,K to 1.08\,K. The results show that QRNN-mwi is partially successful in correcting the cloud impact. A significant reduction in the deviation to NFCS values indicates that though most cases with high cloud impact can be removed/corrected, the cases with residual cloud impact and uncorrected cases deteriorate the performance. Using a low threshold filter can improve the accuracy, but at the cost of excessive removal of clear-sky cases.

@@

High quality cloud corrected radiances have great potential in the DA (data assimilation) systems. Most DA systems reject upto 80\% of the available observations due to cloud contamination. The loss is largest in mid-latitudes. Although all-sky microwave measurements are being successfully assimilated at most operational centres, cloud cleared microwave radiances offer forecast benefits without additional complexities. Assimilation of all-sky radiances calls for a complex radiative transfer and careful consideration of observational errors.
